\chapter{Conclusion and Future Scope}

The development of intelligent fault detection systems for power transmission lines reflects the growing demand for proactive, interpretable, and automated grid management. Addressing this need requires accurate real-time fault classification, system-wide impact analysis, and explainable decision support — capabilities that traditional relay-based and standalone ML systems do not provide together. This chapter summarizes the outcomes of the proposed system and outlines directions for future enhancement.

\section{Conclusion}

This project developed an intelligent real-time fault detection and predictive governance system addressing three key limitations of conventional approaches: delayed fault identification, absence of system-wide impact analysis, and lack of operator decision support.

The system objectives were fully achieved:

\textbf{Objective 1 — Real-time embedded classification:} A Decision Tree model trained on six-dimensional three-phase voltage-current features ($V_a, V_b, V_c, I_a, I_b, I_c$) achieves 96.9\% weighted classification accuracy across all five fault classes (LG, LL, LLG, LLL, No Fault) with 8--12~ms inference latency on the Raspberry Pi, meeting the $>95\%$ accuracy and $<100$~ms latency design targets.

\textbf{Objective 2 — Generative AI integration:} A Llama 3.3-70b model (accessed via the Groq inference API) produces structured root cause explanations, step-by-step mitigation plans, and preventive recommendations for every detected fault event, validated against the injected fault type in all test scenarios.

\textbf{Objective 3 — Sensor-based acquisition:} Three ZMPT101B voltage sensors and three ACS712 current sensors, digitized through an MCP3008 10-bit ADC, provide continuous three-phase measurement with linearity error below 1\% and sampling latency within the 8--12~ms pipeline budget.

\textbf{Objective 4 — Decision Intelligence Engine:} A BFS-based cascade analyzer traverses a 16-node system dependency graph to compute impacted component counts, cascade depth, and a quantitative risk score (0--100) with four-level priority classification (P1 Critical through P4 Low) for every detected event.

\textbf{Objective 5 — Smart Governance Layer:} A policy-driven governance module evaluates risk thresholds in real time and triggers one of three escalation workflows (log-only, admin notification, full mitigation orchestration), with all decisions recorded in a structured ISO-55001-compatible audit log. Governance uptime of 99.2\% was observed over an 8-hour continuous test session.

Compared to existing literature, the proposed system achieves competitive classification accuracy (96.9\% vs.\ 90--97\% for ML-only systems \cite{Zhang2025}) at significantly lower latency (8--12~ms vs.\ 115.2~ms for published edge deep learning \cite{EdgeML2024}), while being the only configuration to integrate cascading impact analysis, LLM reasoning, and governance automation in a single deployable framework.

\section{Future Scope}

The current prototype is validated on a laboratory-scale setup with simulated fault conditions. Several directions exist for extending the system toward industrial deployment:

\textbf{Improved classification accuracy:} Replacing the Decision Tree with a lightweight ensemble (Random Forest with feature subsampling or gradient boosted trees pruned for embedded deployment) could close the gap to the 97--99.96\% accuracy range reported in recent literature \cite{Anwar2025,BPASTS2025} without exceeding Raspberry Pi resource constraints.

\textbf{On-device LLM deployment:} Quantized variants of Llama 3 (4-bit GGUF format) can be deployed on Raspberry Pi using llama.cpp, eliminating the Groq API dependency for offline grid environments. This would reduce AI response latency from 1.8~s to an estimated 5--15~s on the Raspberry Pi, which is acceptable for post-detection analysis.

\textbf{Graph-based topology modeling:} The current dependency graph is static. Integrating a dynamic graph database (e.g., Neo4j) updated from real SCADA telemetry would allow the Decision Intelligence Engine to reflect actual grid topology and handle configuration changes automatically \cite{HybridCascade2025}.

\textbf{Dataset expansion and transfer learning:} Training on larger, publicly available IEEE benchmark fault datasets (e.g., IEEE 14-bus, IEEE 118-bus) and applying transfer learning techniques would improve generalization to diverse grid configurations and operating conditions.

\textbf{IoT and PMU integration:} Replacing the current sensor hardware with Phasor Measurement Units (PMUs) would provide synchronized phasor data across wide-area networks, enabling more accurate cascade modeling and enabling the ensemble XAI methods demonstrated by \textcite{PMUEnsemble2024}.

\textbf{Economic and SLA impact analysis:} Extending the risk scoring model to incorporate economic cost per fault type and SLA breach probability would enable automated self-healing and demand-response decisions, completing the transition from a monitoring system to a fully autonomous grid governance platform.
